### 2020-04-20 Paper/main.tex
Made many minor changes based on feedback

\documentclass[journal]{IEEEtran}

% *** CITATION PACKAGES ***
\usepackage[style=ieee]{biblatex} 
\bibliography{random_walk.bib}

% *** MATH PACKAGES ***
\usepackage{amsmath}

% *** PDF, URL AND HYPERLINK PACKAGES ***
\usepackage{url}
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{graphicx}  % needed to include png, eps figures
\usepackage{float}  % used to fix location of images i.e.\begin{figure}[H]

% *** MY PACKAGES ***
\usepackage{todonotes}
\usepackage[caption=false]{subfig}

\begin{document}

% paper title
\title{Using Random Walk Simulations to Calculate Ground State Energies in
  Quantum Physics}

% author names 
\author{Sai Pandian, ID: 29899923}%

% The report headers
\markboth{PHYS6017 Computer Techniques in Physics Report 1, April 2020}
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
  In this paper, we present a novel way of calculating ground state energies for
  a quantum simulation using a Random Walk Monte Carlo simulation. The random
  walk consists of a walker making steps in random directions, with each step
  having a probability of causing the walker to die and the random walk to end,
  with the probability determined by a function of position. The function can be
  chosen to model specific quantum systems. The rate of death of the walkers is
  used to calculate the ground state energy. This paper explores the efficacy of
  such a method in reproducing the known ground state energy of an infinite
  square well, and the extensibility of the method to more complex
  systems. Using the infinite square well example, the simulation found the
  energy proportionality to be $1.2339 \pm 0.0008$, which is consister to high
  precision with the known $\frac{pi^2}{8}$ theoretical constant. Thus, we found
  that this method is very good at reproducing the known results to a high
  precision, and with good reliability, having very small variance in
  repeats. The method is extensible to higher dimensions, with a linear increase
  in time-complexity for each added dimension.
\end{abstract}

\section{Introduction}
\IEEEPARstart{T}{he} developments in computer power in recent years have opened
up the possibility for novel computational solutions to many problems that were
otherwise limited. We now have access to computers that can numerically solve
problems and run simulations that were previously too computationally intensive
to do.  One such problem is the calculation of the ground state energies of
different quantum systems. Many methods exist for solving these problems, using
both classical and quantum computational approaches. Whilst generally it may be
true that Quantum computating approahes are more suited to solving such problems
\cite{Mazzola}, these approaches are still in infancy and are not readily
accessible. Thus, there exist an abundance of classical computing approaches,
such as the Variational-Relaxiation algorithm \cite{Schroeder2017}. This is a
particularly useful algorithm for two-dimensional systems with non-separable
potentials but quickly becomes difficult to implement for more complex systems.

In this investigation, we explore an alternative Monte Carlo method using Random
Walk simulations, and demonstrate its efficacy with simple systems. This
approach has applications in many areas of Quantum Physics and is values due to
its effectiveness for more complicated systems.

\section{Theoretical Background}
\label{sec:TheoreticalBackground}

\subsection{Random Walk Simulations}

Random walks descibe motion as a series of steps in some vector space, with the
direction of each step being random, i.e. with there being an assigned likelihood
of a step in each possible direction. Consider a simple 1-dimensional line and a
walker standing at a point labelled the origin. The walker takes 20 steps, the
direction of each step (forwards or backwards on the line) is random. The
position of the walker at each step is shown in Figure \ref{fig:cartoon}.

\begin{figure}[H]%[!ht]
  \begin{center}
    \includegraphics[width=0.45\textwidth]{images/cartoon.pdf}
    \caption{The position of a walker at each step in a 20-step random walk
      simulation. The average position of the walker does not necessarily remain
      at the origin as one might expect. Instead it is normal for the
      walker to finish the walk at some distance from the origin, as shown here.}
    \label{fig:cartoon}
  \end{center}
\end{figure}

As we can see from the figure, it is not uncommon for the average position of
the walker to be away from the origin once the walk has concluded. This is
because the position of the walker at any step is dependent only on the position
at the previous step. If the walker happens to take a number of steps in the
same direction, it is highly unlikely for the walker to take an equal number of
steps in the opposite direction, in order to return to the origin. Thus, in such
a case, it is likely that the walker will conclude the walk in a position away
from the origin.

In this example, the walker had an equal likelihood of stepping forward or
backwards, but we can add a bias in a direction if we want to model a particular
phenomenon. The walk is also easily extensible to multiple dimensions, as for
each extra dimension, one extra random number must be drawn. This gives a linear
increase in time-complexity, which is convenient for more complex systems.

\subsection{Application to Quantum Systems}

This investigation applied Random walk simulations to calculate the ground state
energies of Quantum Systems by solving the Schr\"{o}dinger equation, as shown in
\cite{Ceperley1983}.

We can extend the simple one-dimensional line model from Figure
\ref{fig:cartoon} to a quantum system by imagining it as a one-dimensional
simple lattice system. A walker can start at a position labelled the origin, and
then proceed in a random walk. The random walk concludes if the walker ``dies''
on a particular step. The probability that the walker will die is dependent on a
defined potential at that point. For example, in an infinite square well system,
the potential is 0 within a well, and infinite outside the well. So the walker
will have no chance of dying whilst inside the well, but will die if they step
outside.

Let us define $p(j,n+1)$ as the probability that the walker will be at position
$j$ after $n+1$ steps and $a(j)$ as the probability that the walker will die at
position $j$. Thus, for this to be the case, the walker would have had position
$j\pm1$ after $n$ steps. Thus, we can say:

\begin{equation}
  p(j, n+1) =  \frac{1}{2}(1-a(j))(p(j-1,n) + p(j+1,n))
  \nonumber
\end{equation}

If $a(j)$ is small and $n$ is large, we can use Taylor's expansion to expand
each term in this expression and obtain a differential equation. If we then
make an ansatz substitution:

\begin{equation}
  p(j,n) = q(j) \exp(-\lambda n)
  \label{eq:firsteq}
\end{equation}
where $q(j)$ is an arbitrary function and $\lambda$ is the rate of death, the
differential equation in $p(j,n)$ simplifies to:

\begin{equation}
  -\frac{1}{2} \frac{d^2q(j)}{dj^2} + a(j)q(j) = \lambda q(j)
  \label{eq:randomwave}
\end{equation}
The full derivation is taken from \cite{MarcusNewton2020} and is shown in
Appendix \ref{appendix:derivation}. Notice that Equation \ref{eq:randomwave} is
in the form of the Schr\"{o}dinger equation:

\begin{equation}
  \label{eq:schrodinger}
  \frac{-\hbar^2}{2m}\frac{d^2 \psi}{dr^2} + V(r)\psi(r) = E\psi(r)
\end{equation}
Comparing Equation \ref{eq:randomwave} and Equation \ref{eq:schrodinger}, we can
see that the probability of death $a(j)$ is related to the potential $V(r)$,
and the rate of death $\lambda$ is related to the Energy $E$.

We can extend this to an infinite square well by choosing our probability of
death such that:

\begin{equation}
  \label{eq:squarewell}
    a(j) =
    \begin{cases}
      0,& \text{if } j < J\\
      \infty,& \text{if } j \geq J
    \end{cases}
    \nonumber
\end{equation}
where J is a well-defined boundary. So the walker cannot die inside the well but
wil immediately die when they step outside, which we expect will mostly
happen given sufficient time. Inside the well, the potential is simply that of a
free particle \cite{Bransden2003}:

\begin{equation}
  \frac{-\hbar^2}{2m}\frac{d^2 \psi}{dj^2} = E\psi(j)
  \nonumber
\end{equation}
where $x$ is the position of the walker. By making the substitution
$x=J\frac{j}{R}$, such that $j=R$ corresponds to the walker being at the
boundary, we find:

\begin{equation}
  \begin{split}
    -\frac{\hbar^2}{2m}\frac{R^2}{J^2}\frac{d^2\psi}{dj^2} &= E\psi(x)\\
    -\frac{1}{2}\frac{d^2\psi}{dj^2} &= \frac{mJ^2E}{\hbar^2R^2}\psi(x)
  \end{split}
  \nonumber
\end{equation}
If we compare this to Equation \ref{eq:randomwave}, allowing $a(j) = 0$ inside
the well, we see that:

\begin{equation}
  E = \lambda R^2 \frac{\hbar^2}{mJ^2}
  \label{eq:kasia}
\end{equation}
where only $\lambda$ is an unknown quantity. Thus, by calculating rate of death
$\lambda$, we can calculate the ground state energy of the infinite square
well. The known ground state energy of an infnite square well is
$\frac{\pi^2}{8}\frac{\hbar^2}{mJ^2}$ \cite{Davies1998}, so we expect to find $\lambda
R^2$ close to $\frac{\pi^2}{8}$.

To find the rate of death $\lambda$, we consider Equation \ref{eq:firsteq}. With
some simple manipulation:
\begin{equation}
  \begin{split}
    p(j, n) & = q(j) \exp(-\lambda n)\\
    \log(p) & = \log(q \exp(-\lambda n)) \\
  \end{split}
  \nonumber
\end{equation}

\begin{equation}
  \log(p) = -\lambda n + log(q)
  \label{eq:straightline}
\end{equation}

Thus, by plotting the logarithm of the probability of surviving $n$ steps
against the number of steps $n$, we can obtain the death rate $\lambda$ from the
gradient of the line.

\section{Method}
We implemented the Random Walk simulation computationally in
C++ in order to optimise the performance of the simulation, and thereby allow
for a larger number of iterations and larger potential wells. The source code is
available \cite{Pandian}. The simulation is easily extensible to multiple
dimensions with a linear increase in time-complexity as the number of dimensions
increases, but this investigation focused on a one-dimensional system.

The random direction is selected using a pseudorandom number generator (pRNG)
known as the Mersenne Twister. This algorithm was chosen as it is a very fast
implementation of a pRNG \cite{Route2017}, passes many statistical randomness
tests \cite{Bellamy2013}, and has a very large period of $2^{19937}-1$.

We initially ran the simulation for a boundary of $J = 20$, with $1000000$
walkers. The step on which each walker died was recorded so that a probability
distribution showing the likelihood of dying on a particular step could be
constructed.

We then considered the probability that a walker will survive $n$ steps, and use
the relationship between the number of walkers surviving $n$ steps and $n$ to
determine the rate of death $\lambda$, as shown in Equation \ref{eq:straightline}.

One method of reducing the computing time of the simulation is to introduce a
maximum number of steps. The walker will die if they reach the boundary of the
well, or if they reach this maximum number of steps, whichever happens
first. The effect of this method on the result is investigated in this paper.

% However, this will introduce some systematic error into the simulation,
% as there will be a disproportionately high number of walkers dying on the
% maximum number of steps. We also investigated these effects here.

% We expect to see an effect on the accuracy of our final ground state depending
% on the size of the infinite well. We investigate these effects by comparing the
% calculated ground state energy to the theoretical energy for different sized wells.

\section{Results and Discussion}

\begin{figure*}[ht!]
  \centering
  \subfloat[Linear Scale]
  {\includegraphics[width=0.5\textwidth]{images/cum_exp_plot.pdf}
  \label{fig:cumhist}}
  \centering
  \subfloat[Log Scale]
  {\includegraphics[width=0.5\textwidth]{images/cum_line_plot.pdf}
  \label{fig:cumloghist}}
  \caption{Histograms showing the Probability of surviving to at least a
    particular step in an infinite square well with $J = 20$. The peak of 1 does
    not start at the origin but at step 100. This is because these steps were
    neglected in the plot as prior to this step the exponential shape is less
    clear. An exponential decay curve provides a good fit, with an $R^2$ value
    of 0.9999998. Plotting on a logarithmic scale shows the straight line fit,
    the gradient of which is $\lambda$, giving $\lambda J^2 = 1.2334$, with a
    negligible error. The fluctuation in the higher step number regions is due
    to the large variance in the small number of walkers that die at these
    steps.}
  \label{fig:cumplots}
\end{figure*}

% The simulation was first run for an infinite square well with boundary $J=20$
% with 1000000 walkers. The number of walkers surviving at each step was recorded
% and the histograms in Figure \ref{fig:cumplots} were created. The probability
% that a walker will survive at a particular step is simply the 1 minus the
% normalised cumulative sum of walkers dying at each step. The curve has been
% fitted with an exponential function of the form in Equation
% \ref{eq:firsteq}.

We first investigated an infinite square well with boundary $J=20$ and 1000000
walkers. We found the energy constant to be $1.2339 \pm 0.0008$, calculated from
the probability distribution of survival at a particular step. The probability
that a walker will survive at a particular step is simply the 1 minus the
normalised cumulative sum of walkers dying at each step. Figure
\ref{fig:cumplots} presents this probability and confirms the Anstatz of the
exponential function. The curve has been fitted with an exponential function of
the form in Equation \ref{eq:firsteq}. The gradient of Figure
\ref{fig:cumloghist} allowed for calculation of death rate which from Equation
\ref{eq:kasia} gives the calculated energy.

Figure \ref{fig:cumplots} results in a very good fit, with an $R^2$ value of
0.9999998, indicating the data does not vary much from the curve of best
fit. This is likely due to the very large number of walkers used in this
experiment, as allowed by the performant C++ implementation.  We can see this
better in Figure \ref{fig:cumloghist}, which is the same exponential plot fitted
on a logarithmic scale. The closeness of the data to the line of best fit is
evident here. The error in the fitted parameters is negligible, as shown by the
$R^2$ value. The gradient of the fitted line is taken to be $\lambda$ and thus
allows us to calculate $\lambda J^2 = 1.2334$, compared to the theoretical value
of $\frac{\pi^2}{8} = 1.2337$, quoting both values to the precision at which
they vary. Since the numerical error in the fits is negligible, it cannot be
propagated into an error in the final answer. Thus, we get the error in the
final answer from repeated runs of the simulation and taking the standard
deviation of the many values.

As we can observe in Figure \ref{fig:cumhist}, the peak of the curve does not
lie at the origin as we might expect from the fact that all walkers should
survive at least the first step. The reason for this is twofold:
\begin{itemize}
\item For a well with boundary $J$, it is not possible for the walker to die in
  a number of steps less than $J$
\item It is highly unlikely that the walker will die at a number of steps
  greater than but still close to $J$ since it would require walking in the same
  direction a dispropotionate number of times.
\end{itemize}

\begin{figure}[H]%[!ht]
  \begin{center}
    \includegraphics[width=0.45\textwidth]{images/exp_plot.pdf}
    \caption{The probability of dying on a particular step in an infnite square
      well with $J=20$. As can be seen, the peak is away from the origin. This
      is because, for a well with boundary $J$, the minimum number of steps a
      walker must take is $J$. However, it is still unlikely that the walker
      will die on this minimum number of steps, as then the walker must walk in
      the same direction every time. Thus, the most probable step for the walker
      to die on is greater than $J$. It is for this reason that the exponential
      decay does not begin at the origin in Figure \ref{fig:cumhist}, meaning
      some earlier steps must be neglected in order to achieve a better fit.}
    \label{fig:hist}
  \end{center}
\end{figure}

These effects are evident in Figure \ref{fig:hist}, which presents the
probability that a walker will die at a particular number of steps. We can see
that the peak of the curve is very clearly at a step that is greater than 20,
which was the $J$ value used in this simulation. Thus, for low step numbers, the
shape of the curve is not exponential. For this reason we used a simple
algorithm to identify a suitable number of steps to skip before fitting the
exponential curve. It is for this reason that the histogram in Figure
\ref{fig:cumhist} does not start at the origin.

\begin{figure}%[!ht]
  \begin{center}
    \includegraphics[width=0.45\textwidth]{images/line_plot.pdf}
    \caption{The probability of dying on a particular step in an infnite square
      well with $J=20$ as in Figure \ref{fig:hist}, plotted on a logarithmic
      scale. As explained for Figure \ref{fig:hist}, the peak probability is on
      a step greater than $J$. This is clear in this plot. It is also evident
      that there is significantly more spread in the points at higher
      steps. This is due to there being significantly fewer walkers dying on
      these steps, so the relative differences are higher. This explains the
      fluctuations seen at higher steps in Figure \ref{fig:cumloghist}.}
    \label{fig:loghist}
  \end{center}
\end{figure}

We observe some fluctuation in the data in the higher steps in Figure
\ref{fig:cumloghist}. This is because there are so few walkers surviving to this
many steps that there is more significant variance in their relative numbers. We
can observe this in Figure \ref{fig:loghist}. At higher step numbers, the spread
is significantly larger due to the greater relative variance in the small number
of walkers surviving to this many steps. When this data is cumulatively summed
to calculate the probability of survival shown in Figure \ref{fig:cumloghist},
the spread is present as the fluctuations.

\begin{figure*}[ht!]
  \centering
  \subfloat[Probability of dying on a particlar step in Linear scale]
  {\includegraphics[width=0.5\textwidth]{images/exp_plot_cutoff.pdf}
  \label{fig:expcutoff}}
  \centering
  \subfloat[Probability of surviving to at least a particular step in a
  logarithmic scale]
  {\includegraphics[width=0.5\textwidth]{images/cum_line_plot_cutoff.pdf}
  \label{fig:cum_line_cutoff_plot}}
  \caption{Histograms showing the results of the simulation when a maximm number
  of steps is introduced. If a walker reaches this number of steps, they
  automatically die, regardless of position. This results in a
  disproportionately high number of walkers dying at this maximum number, 2000
  in this case, as shown by Figure \ref{fig:expcutoff}. Although this presents
  no fluctuation at higher number of steps, as shown in Figure
  \ref{fig:cum_line_cutoff_plot}, some systematic error is introduced
  through this method.}
  \label{fig:cutoffplots}
\end{figure*}

It is possible to remove this variance by introducing a cutoff number of steps,
i.e. a maximum number of steps at which the walker will automatically die,
regardless of whether they reached a boundary or not. Thus, there will not be
such high variance in the number of walkers surviving to high number of
steps. This is presented in Figure \ref{fig:cum_line_cutoff_plot}. While the
Figure \ref{fig:cum_line_cutoff_plot} fit is observed to be better, with a
higher $R^2$ value, some systematic error has been introduced since there will
now be a disproportionately high number of walkers dying at the cutoff step as
can be seen in Figure \ref{fig:expcutoff}. Thus, the final $\lambda J^2$ value
will be less accurate. On a faster computer, with more efficient code, it is
possible to forgo the step cutoff and allow the simulation to run until all the
walkers die. This may take much longer but will ensure there will not be such a
systematic error introduced. This is the method used in this investigation.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=0.45\textwidth]{images/multiplot.pdf}
    \caption{Histogram plotted on a logarithmic scale showing the Probability of
      surviving to at least a particular step in different infinite square wells
      with different boundaries $J$. The simulations were run with $10000$
      walkers The $\lambda J^2$ values for each are shown. The values get closer
      to the theoretical value of $\frac{\pi^2}{8} = 1.2337$ as we increase the
      size of the well. This is likely due to more walkers surviving to a higher
      number of steps, as the well size is increased.}
    \label{fig:multi_line_plot}
  \end{center}
\end{figure}

It is also necessary to investigate how varying the boundary $J$ affects the
accuracy of the final answer. Figure \ref{fig:multi_line_plot} presents the
probability plot from Figure \ref{fig:cumloghist} for many different boundary
$J$ values. No step cutoff was introduced. The calculated $\lambda J^2$ value is
shown in the legend. The gradient evidently changes, since the expected highest
number of steps would increase as the boundary is larger, as there is more
space for the walker to move before reaching a boundary. As shown in the legend,
the calculated $\lambda J^2$ values get closer and closer to the
theoretical value as $J$ increases. This is due to the fact that as $J$
increases, there are more walkers surviving to larger number of steps. Thus, we
can say that using a larger value of $J$ will result in a more accurate answer
but will also be more computationally intensive and take longer. Thus, we make a
tradeoff between speed and accuracy and choose $J=20$ for the final
investigation.

Running the simulation 10 times with 1000000 walkers in an infinite square well
with $J=20$ and taking the average $\lambda J^2$ with error being standard
deviation, we find:
\begin{equation}
  \lambda J^2 = 1.2339 \pm 0.0008
  \nonumber
\end{equation}
We quote the precision to the decimal place at which there is variation from the
theoretical result of $1.2337$. We notice that the error in the calculated
answer is small (since there is not much variation between repeated simulations)
and the theoretical result is within the range of the calculated answer.

\section{Conclusions}

In this investigation we demonstrated a method of calculating ground state
energies of a quantum system using Random walk computational simulations. The
investigation focussed on a one dimensional infinite square well, but can easily
be extended to more complex systems such as a cirular well in two dimensions.

The investigation explored the effects of varying the size of the well in
determining the ground state energy of a quantum system, and showed that this
Monte Carlo method produces more accurate results for ground state energies of
larger wells.  This is because for wells with larger boundaries, the walker can
take a larger number of steps on average before dying, and thus can populate
larger step bins. Thus, this method produces data that is close to the
theoretical result for larger infinite square wells.

The numerical result found the ground state energy proportionality constant to
be $\lambda J^2 = 1.2339 \pm 0.0008$, which is consistent with the known
theoretical value of $\frac{\pi^2}{8}$.

This method can be easily extended to multiple dimensions with only a linear
increase in time complexity for each added dimension. Thus, while this method
may not be very efficient for low number of dimensions, it quickly becomes
effective at a higher number of dimensions.

\pagebreak
\appendices
\section{Random walk Schr\"{o}dinger Equation Derivation}
\label{appendix:derivation}
Consider a one-dimensional lattice upon which a walker is undergoing a random
walker. Let us denote the probability that the walker is at position $j$ after
$n$ steps as $p(j,n)$ and the probabilty that the walker will die at position
$j$ as $a(j)$. We will denote the position-step number coordinates as $[j,
n]$. Thus, for a walker to have coordinates $[j, n+1]$, in the previous step,
they must have had coordinates $[j \pm 1, n]$. So we can say:

\begin{equation}
  \label{eq:probability}
  p(j, n+1) =  \frac{1}{2}(1-a(j))(p(j-1,n) + p(j+1,n))
\end{equation}

If $a(j)$ is small and $n$ is large, then we can say $p(j, n)$ is a continuous
function and we can use Taylor expansion to expand each of $p(j+1, n)$,
$p(j-1,n)$, and $p(j, n+1)$:

\begin{equation}
  p(j+1, n) = p(j,n) + \frac{\partial p}{\partial j} + \frac{1}{2}
  \frac{\partial^2 p}{\partial j^2} + ...
  \nonumber
\end{equation}
\begin{equation}
  p(j-1, n) = p(j,n) - \frac{\partial p}{\partial j} + \frac{1}{2}
  \frac{\partial^2 p}{+\partial j^2} + ...
  \nonumber
\end{equation}
\begin{equation}
  p(j, n+1) = p(j, n) + \frac{\partial p}{\partial n} + ...
  \nonumber
\end{equation}

We can now substitute these expansions into Equation \ref{eq:probability}, we
find:

\begin{equation}
  p(j, n) + \frac{\partial p}{\partial n} = (1-a(j))\Big(p +
  \frac{1}{2}\frac{\partial^2 p}{\partial j^2}\Big)
  \nonumber
\end{equation}

But the probability of death $a(j)$ and the second derivative are so small, that
their product is neglected. We also let $p(j,n) = q(j) \exp(-\lambda n)$ so that:

\begin{equation}
  -\frac{1}{2} \frac{d^2q(j)}{dj^2} + a(j)q(j) = \lambda q(j)
  \nonumber
\end{equation}

\printbibliography

\end{document}